Consistent Polyhedral Surrogates via Embedding

### 2021-05-12 notes/jessie-russian-lemma.tex
Updates on top k and selection

\title{Lemma 1}
\author{Jessie}

@@

Recall the set $\Z_u(p) = \{\E_p V \mid V \in \V_u\}$.
\begin{lemma}
	The set $\Z_u(p) \subseteq \reals^d$ is convex and closed.
\end{lemma}
\begin{proof}
	Recall that $V \in \V_u$ implies that $V(\omega) \in \partial L(u,\omega)$ for all $\omega \in \Omega$ and $V$ measurable.

@@

\section{More direct proof}
Consider the sets
\begin{align*}
A &:= \partial \E_p L(u,Y)\\
B &:= \{\E_p V(Y) \mid V\textrm{ measurable, } V(y) \in \partial L(u,y) \, p-a.s.\}\\
Z &:= \{\E_p V(Y) \mid V\textrm{ measurable, } V(y) \in \partial L(u,y) \, \forall y \in \Y \} 
\end{align*}

We want to show $A = Z$; the Russian paper shows $A =B$.
First, we have $A \supseteq B$ by definition of subdifferential, and $B \supseteq Z$ trivially, so $A \supseteq Z$.
It is left to show $A \subseteq Z$.

As $A$ is a subdifferential of a finite, convex loss defined on $\reals^d$, we have $A$ convex and compact.

In particular, if \jessie{CONDITION} in $A$ are described by elements of $Z$, then anything in $A$ can be described by an element of $Z$ by its convexity and closedness.

Hypothesis: \jessie{CONDITION} = exposed faces in the relative boundary of $A$.

@@

\jessie{This isn't quite typechecking for me, with the function $P$.  In their paper, $P$ should be a function of a subset of $\Y$, but for monotonicity, we want it to be the CDF?}

If we want to show $A := \{\E_p V \mid V(y) \in \partial L(u,y) \, p-a.s.\}$ is closed, and do this by showing $A = cl(A)$.  Thankfully, $A \subseteq cl(A)$ follows trivially, so it is left to show $A \supseteq cl(A)$.

Consider a sequence $\{v_m\} \to v$ such that $v_m \in A$ for all $m$.  We have (by part 1 of the lemma) measurable selections $V_m$ such that $v_m = \E_p V_m$ and the $v_m$ has bounded norm.
If we have $v \in A$, then we have $A \supseteq cl(A)$, and therefore $A$ closed.

Now define functions $P_m : 2^\Y \to \reals^d$ so that $P_m(E) = \int_E V_m(y) dpy$.  
We can use Helly's Theorem (Theorem~\ref{thm:helly}?) to claim that $P_m$ converges pointwise to some function $P$ of bounded variation.\jessie{We need montonicity of the functions $P_m$, but is this definition $A \subseteq B \implies f(A) \leq f(B)$? We don't have that.}
We have uniform boundedness from their first past of the result, and monotonicity from the nonnegativity of probability densities. \jessie{Nope}%from each $P_m$ being a CDF.

\begin{theorem}[Helly's Theorem]\label{thm:helly}
	A uniformly bounded sequence of monotone real functions admits a (pointwise) convergent subsequence
\end{theorem}

Therefore, for any $z \in \reals^d$, by the Dominated Convergence Theorem \jessie{or Helly-Bray \url{https://en.wikipedia.org/wiki/Helly\%E2\%80\%93Bray_theorem}}, we have 
\begin{align*}
\int \inprod{z}{P(y)} dy &= \lim_{m \to \infty} \int \inprod{z}{P_m(y)} dy
\end{align*}

In particular, we can now take the selection $V(y) = P(y)$ as the density of $P$.
Define $P_c$ as the singular part of $P$; that is, the region where $P$ is not defined, concentrated on a set of measure zero.

If $P_c$ is empty (i.e., $P$ is defined for all $y \in \Y$)\jessie{I think this is what we need to claim for an everywhere selection}, then for any $z \in \reals^d$, we have
\begin{align*}
\int \inprod{z}{P(y)} dy &= \lim_{m \to \infty} \int \inprod{z}{P_m(y)} dy \leq \int L'(u,y;z) dy
\end{align*}

From this, it follows that $\inprod{z}{V(y)} \leq L'(u,y;z)$ for all $y \in \Y$, and therefore $V(y) \in \partial L(u,y)$ for all $y \in \Y$.

This yields $P(\Y) = \int_\Y V(y) dpy \in \int_\Y \partial L(u,y) dpy$.

Therefore, we have 
\begin{align*}
\lim_{m\to \infty} V_m &= V \\
\implies \lim_{m \to \infty} \E_p V_m &= \E_p V \in A\\
\lim_{m \to \infty} v_m &= v \in A
\end{align*}

@@

Therefore, we can take $V'$ to be any function that matches $V$ on $y \in E$, and takes any $V'(y) \in \partial L(u,y)$ (which is nonempty and compact by minnability and finiteness of $L$, respectively) elsewhere. \jessiet{Not sure if we want something explicit, but wasn't sure how to define it to be unique.  Mostly was thinking about minning the norm, but still needs more specificity.}
As $V$ is measurable and $p$ is a complete measure, we have $V'$ measurable.
Consider that for all $x \in \reals^d$, we have $V'^{-1}(x) = V^{-1}(x) \cup \{y \in E^c : V'(y) = x\}$.
As $V$ is measurable, we have $V^{-1}(x) \in B$ and since $p$ is a complete measure, we know that $\{y \in E^c : V'(y) = x\} \in B$ as well.
Since sigma algebras are closed under finite unions, we have $V'^{-1}(x) \in B$ for all $x$, and therefore $V'$ is measurable.

>>>>

\title{Generalizing the result of Ioffe and Tikhomirov}
\author{Jessie}

@@

Recall the set $\Z_u(p) = \{\E_p V \mid V \in \V_u\}$.
\begin{lemma}\label{lem:Z-convex-closed}
	The set $\Z_u(p) \subseteq \reals^d$ is convex and closed.
\end{lemma}
\begin{proof}
	Recall that $V \in \V_u$ implies that $V(\omega) \in \partial L(u,\omega)$ for all $\omega \in \Omega$ and $V$ measurable.

@@

%\section{More direct proof}
%Consider the sets
%\begin{align*}
%A &:= \partial \E_p L(u,Y)\\
%B &:= \{\E_p V(Y) \mid V\textrm{ measurable, } V(y) \in \partial L(u,y) \, p-a.s.\}\\
%Z &:= \{\E_p V(Y) \mid V\textrm{ measurable, } V(y) \in \partial L(u,y) \, \forall y \in \Y \} 
%\end{align*}
%
%We want to show $A = Z$; the Russian paper shows $A =B$.
%First, we have $A \supseteq B$ by definition of subdifferential, and $B \supseteq Z$ trivially, so $A \supseteq Z$.
%It is left to show $A \subseteq Z$.
%
%As $A$ is a subdifferential of a finite, convex loss defined on $\reals^d$, we have $A$ convex and compact.
%
%In particular, if \jessie{CONDITION} in $A$ are described by elements of $Z$, then anything in $A$ can be described by an element of $Z$ by its convexity and closedness.
%
%Hypothesis: \jessie{CONDITION} = exposed faces in the relative boundary of $A$.

@@

\jessie{their lemma}
\begin{lemma}
	The set $\E_q \partial L(u,Y) = \{\E_q V \mid V(y) \in \partial L(u,y) \, q-a.s.\} =: A$ is convex and closed.
\end{lemma}

\jessie{What we want to show}
\begin{lemma}
	The set $Z = \{\E_q V \mid V \in \V_u\}$ is convex and closed.\jessiet{$Z$ and $\Z_u(q)$ are the same thing, but one is easier to type.}
\end{lemma}

For now, let's take the convexity of $Z$ for granted.  
I think it's in lemma~\ref{lem:Z-convex-closed}.
If we want to show $Z$ is closed, and do this by showing $Z = cl(Z)$.  Thankfully, $Z \subseteq cl(Z)$ follows trivially, so it is left to show $Z \supseteq cl(Z)$.

Consider a sequence $\{v_m\} \to v$ such that $v_m \in Z$ for all $m$.  We have (by part 1 of the lemma) measurable selections $V_m$ such that $v_m = \E_p V_m$ and the $v_m$ has bounded norm.
If $v \in Z$, then we have $Z \supseteq cl(Z)$, and therefore $Z$ is closed.

Now define functions $p_m : 2^\Y \to \reals^d$ so that $p_m(E) = \int_E V_m(y) dpy$.  
We can use Helly's \jessie{Selection?} Theorem (Theorem~\ref{thm:helly}) to claim that $p_m$ converges pointwise to some function $p$ of bounded variation.
We have uniform boundedness from their first past of the result, and monotonicity from the nonnegativity of probability densities.

\begin{theorem}[Helly's Selection Theorem]\label{thm:helly}
	Let $f_n:\Y \to \reals$ be a sequence of functions with uniformly bounded total variation, and $(f_n)$ is uniformly bounded at a point.
	Then there exists a convergent subsequence $f_{n_k}$ of $f_n$ and function $f$ of bounded variation such that $f_{n_k} \to f$ pointwise.
\end{theorem}

Therefore, for any $z \in \reals^d$, by the Dominated Convergence Theorem, we have 
\begin{align*}
\int \inprod{z}{p(\{y\})} dqy &= \lim_{m \to \infty} \int \inprod{z}{p_m(\{y\})} dqy
\end{align*}

In particular, we can now take the selection $V(y) = p(\{y\})$ as the density of $p$.
(Note: this implies $p(\Y) = \E_q V$.)

We have $p(\{y\})$ is defined for all $y \in \Y$ by pointwise convergence in Theorem~\ref{thm:helly}\jessiet{??}, and therefore we can take, for any $z \in \reals^d$, 
\begin{align*}
\int \inprod{z}{p(\{y\})} dqy &= \lim_{m \to \infty} \int \inprod{z}{p_m(\{y\})} dqy \leq \int L'(u,y;z) dqy
\end{align*}

From this, it follows that $\inprod{z}{V(y)} \leq L'(u,y;z)$, and therefore $V(y) \in \partial L(u,y)$, for all $y \in \Y$.

This yields $p(\Y) = \int_\Y V(y) dqy = \E_q V \in Z$.

Therefore, we have 
\begin{align*}
\lim_{m\to \infty} V_m &= V \\
\implies \lim_{m \to \infty} \E_q V_m &= \E_q V \in Z\\
\lim_{m \to \infty} v_m &= v \in Z
\end{align*}
Thus, $v \in Z$, and $Z$ is closed.

\subsection{After the Lemmas...}
With the Lemmas in hand, the proof is completed by contradiction.
The task is to show $\partial \E_p L(u,Y) \subseteq \E_p \partial L(u,Y)$.

Suppose there is a $x \in \partial \E_p L(u,Y)$ such that $x \not \in \E_p \partial L(u,Y)$.

Then there is a hyperplane in direction $z$ strictly separating $x$ from $\E_p \partial L(u,Y)$.

By Lemma 1, we know that there is some $x' \in \E_p \partial L(u,Y)$ so that $\inprod{z}{x'} = (\E_p L)'(u,Y;z)$, and it follows that this directional derivative can be strictly separated from $x$.
Therefore, we have $x \not \in \partial \E_p L(u,Y)$, contradicting our original assumption.

@@

Therefore, we can take $V'$ to be any function that matches $V$ on $y \in E$, and takes any $V'(y) \in \partial L(u,y)$ (which is nonempty and compact by minnability and finiteness of $L$, respectively) elsewhere. \jessiet{Not sure if we want something explicit, but wasn't sure how to define it to be unique.  Mostly was thinking about minning the norm, but still needs more specificity.}
As $V$ is measurable and $p$ is a complete measure, we have $V'$ measurable.
Consider that for all $X \in \B(\reals^d)$, we have $V'^{-1}(X) = V^{-1}(X) \cup \{y \in E^c : V'(y) \in X\}$.
As $V$ is measurable, we have $V^{-1}(X) \in B$ and since $p$ is a complete measure, we know that $\{y \in E^c : V'(y) \in X \} \in B$ as well.
Since sigma algebras are closed under finite unions, we have $V'^{-1}(X) \in B$ for all $X \in \B(\reals^d)$, and therefore $V'$ is $B$-measurable.

### 2021-05-12 notes/top-k.tex
Updates on top k and selection

\begin{lemma}\label{lem:linear-pieces}
	For each element $u_i$ of $u \in \U$, the expected loss $\inprod{p}{L^{top-k}(u)}$ is linear on $[0,1]$ and $[1, \bar u_{-i} + 1]$ for all $p \in \simplex$.
	\jessie{Not sure if this is the right statement.}
    \raf{Not sure either.  One thing to be careful about: you either have to make sure that the intervals in question yield points in the set $\U$, or refer to the original form of the loss.  Because the form of the loss you are using only is shown to hold when $u\in\U$.  Hopefully that made sense.}
    \raf{One thing that would help: let's visualize $\U$ for some value of $k,n$ to get some intuition for what this set is.  Maybe $k=2$ and $n=3$, but only plot the first 2 coordinates?}
\end{lemma}
\begin{proof}
	For all $p \in \simplex$ and $u \in \U^\circ$, we know that 
	\begin{equation}
	\inprod{p}{L^{top-k}(u)} = 1 - \bar u + \inprod{p}{u + \frac 1 k \min{u,\ones}}~.~
	\end{equation}
	For any $p$, if $u_i \in [0,1]$, we then have 
	\begin{align*}
	\inprod{p}{L^{top-k}(u)} &= 1 + \bar u - \inprod{p}{u + \frac 1 k \min{u,\ones}}\\
	 &= 1 + \bar u - \sum_{y \neq i}p_y (u_y + \frac 1 k \min(u_y, 1)) - p_i \frac {k+1} k u_i~,~
	\end{align*}
	which is linear on $u_i$.

@@

\begin{theorem}\label{thm:finite-set}
	Consider the loss $L^{top-k} : \reals^n \times \Y \to \reals_+$ given in Equation~\ref{eq:top-k-surrogate}.  
	For all $p \in \simplex$, there is a report in the following report set $\R$ minimizing $\inprod{p}{L^{top-k}(\cdot)}$.
	\begin{align*}
	\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}
	\end{align*}
\end{theorem}
\begin{proof}[Proof Sketch.]
%	The above work tells us that when $p_y > \frac 1 {k+1}$, it is optimal to set $u_y > 0$, and otherwise, it is optimal to set $u_y = 0$.
	Since $L^{top-k}$ is linear on $[0,1]$ and $[1, c]$ for some $c \geq 1$ and all dimensions $i \in [n]$, we know that the optimum is on the boundary of such reports, so when we bump $u_y$ up to some number greater than $0$, it should either be one $1$ or $c$ to be an optimal report.
	$H$ represents the elements of the report that are bumped up to $c$, and $M$ represents the elements of the report that are bumped up to $1$.
	We know that the size of $H \cup M$ must be less than or equal to $k$ by Assumption~\ref{assum:nonneg}, and $H \cap M$ is empty by each element of $u$ being real-valued, so for $c > 1$, $u_i$ cannot simultaneously be equal to $c$ and $1$, hence the restrictions on $\R$.
	Note that if $c = 1$, then we can say $H = \emptyset$.

@@

>>>>

\begin{lemma}\label{lem:linear-pieces}
	For each element $u_i$ of $u \in \U$, the expected loss $\inprod{p}{L^{top-k}(u)}$ is linear on $[0,1]$ and $[1, \bar u_{-i} + 1]$ for all $p \in \simplex$.
	\jessie{Not sure if this is the right statement.}
    \raf{Not sure either.  One thing to be careful about: you either have to make sure that the intervals in question yield points in the set $\U$, or refer to the original form of the loss.  Because the form of the loss you are using only is shown to hold when $u\in\U$.  Hopefully that made sense.}
    \raf{One thing that would help: let's visualize $\U$ for some value of $k,n$ to get some intuition for what this set is.  Maybe $k=2$ and $n=3$, but only plot the first 2 coordinates?}
\end{lemma}
\begin{proof}
	For all $p \in \simplex$ and $u \in \U^\circ$, we know that 
	\begin{equation}
	\inprod{p}{L^{top-k}(u)} = 1 + \bar u - \inprod{p}{u + \frac 1 k \min{u,\ones}}~.~
	\end{equation}
	For any $p$, if $u_i \in [0,1]$, we then have 
	\begin{align*}
	\inprod{p}{L^{top-k}(u)} &= 1 + \bar u - \inprod{p}{u + \frac 1 k \min(u,\ones)}\\
	 &= 1 + \bar u - \sum_{y \neq i}p_y (u_y + \frac 1 k \min(u_y, 1)) - p_i \frac {k+1} k u_i~,~
	\end{align*}
	which is linear on $u_i$.

@@

\begin{theorem}\label{thm:finite-set}
	Consider the loss $L^{top-k} : \reals^n \times \Y \to \reals_+$ given in Equation~\ref{eq:top-k-surrogate}.  
	The finite set
	\begin{align*}
	\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}
	\end{align*}
	is representative for $L^{top-k}$.
\end{theorem}
\begin{proof}[Proof Sketch.]
%	The above work tells us that when $p_y > \frac 1 {k+1}$, it is optimal to set $u_y > 0$, and otherwise, it is optimal to set $u_y = 0$.
	Since $L^{top-k}$ is linear in $i$ on $[0,1]$ and $[1, c]$ for some $c \geq 1$ and all dimensions $i \in [n]$, we know that the optimum is on the boundary of such reports, so when we bump $u_y$ up to some number greater than $0$, it should either be one $1$ or $c$ to be an optimal report.
	$H$ represents the elements of the report that are bumped up to $c$, and $M$ represents the elements of the report that are bumped up to $1$.
	We know that the size of $H \cup M$ must be less than or equal to $k$ by Assumption~\ref{assum:nonneg}, and $H \cap M$ is empty by each element of $u$ being real-valued, so for $c > 1$, $u_i$ cannot simultaneously be equal to $c$ and $1$, hence the restrictions on $\R$.
	Note that if $c = 1$, then we can say $H = \emptyset$.

@@

\section{Property elicited by $L^{top-k}$}
Consider that we have $\R$ as above, but we are not sure when to ``bump'' reports up or down from one group to another.
There is sometimes a benefit in expected loss to artificially bumping up from the ``low'' to ``middle'' bin when possible because doing so cranks up the constant on the ``high'' reports, yielding better expected loss.

Note that ``bumping down'' does not happen because of how the original bins are set, but we can ''bump up'' from low to medium or medium to high.

Given $p \in \simplex$, we have the original bins
\begin{align*}
\bar H(p) &:= \{i \mid p_i \geq \frac 1 k\}\\
\bar M(p) &:= \{i \mid p_i \in [\frac 1 {k+1}, \frac 1 k]\} \\
\bar L(p) &:= \{i \mid p_i \leq \frac 1 {k+1}\}~,~
\end{align*}
dropping $p$ from the notation where it can be understood from context.

Now we want to modify this bins so that we can ``bump'' reports up to a higher bin if beneficial.
One can verify that it is never beneficial to bump from middle to high, so we need only concern ourselves with bumping from $L$ to $M$.
To understand when to bump $j$ from lower to middle, taking $u \to u'$, we can ask when we have
\begin{align*}
\inprod{p}{L(u)} &\geq \inprod{p}{L(u')}\\
\sum_{i \in \bar H} \frac{|\bar M| + k -1}{k-|\bar H|} - p_i \left(\frac{k(|\bar M| + k -1)+ 1}{k(k- |\bar H|)}\right) + \sum_{i \in \bar M} \left(\frac 1 k - p_i \left( \frac {k+1}{k}\right)\right) &\geq \sum_{i \in \bar H} \frac{|\bar M| + k }{k-|\bar H|} - p_i \left(\frac{k(|\bar M| + k)+ 1}{k(k- |\bar H|)}\right) + \sum_{i \in \bar M'} \left(\frac 1 k - p_i \left( \frac {k+1}{k}\right)\right)\\
\frac{|\bar H|(|\bar M| + k - 1)}{k(k- |\bar H|)} + p_H \frac{-k(|\bar M| + k - 1) - 1}{k(k - |\bar H|)} &\geq \frac{|\bar H|(|\bar M| + k)}{k(k - |\bar H)} + p_H \left(\frac{-k(|\bar M| + k)-1}{k (k - |\bar H|)}\right) + \left(\frac 1 k - p_j \left(\frac {k+1}{k}\right)\right) \\
\frac{-|\bar H|}{k (k - |\bar H|)} + \frac{p_H}{k - |\bar H|} &\geq \frac 1 k - p_j \left(\frac{k+1}{k}\right)~,~
\end{align*}
where $p_H := \sum_{i \in \bar H} p_i$.
Thus, we can take the set 
\begin{align}
B(p) &:= \left\{i : p_i \leq \frac{1}{k+1} \wedge \frac{p_H k- |\bar H|}{k(k- |\bar H|)} \geq \frac 1 k - p_i\left(\frac{k+1}{k}\right) \right\}
\end{align}

Now take the sets
\begin{align*}
H(p) &:= \bar H(p) \\
M(p) &:= \bar H(p) \cup B(p) \\
L(p) &:= \bar L(p) \setminus B(p)\\
\end{align*}
and take the function 
\begin{align}
\phi(p) := \frac{|M(p)| + k - 1}{k-|H(p)|} \ones_{H(p)} + \ones_{M(p)}~.~
\end{align}
\jessie{Fix with algorithm}

\begin{proposition}
	For all $p \in \simplex$, we have $\phi(p) \in \Gamma(p) \cap \R$.
\end{proposition}
\begin{proof}
	To see $\phi(p) \in \R$ by construction. 
\end{proof}

\begin{corollary}
	$\phi(\simplex)$ is a finite representative set.
\end{corollary}
